HEP Software Foundation Community White Paper Working Group -- Training, Staffing and Careers

### 2018-06-11 CWP/papers/HSF-CWP-2017-02_training/latex/cwp-wg-training-careers.tex
Restructure paper to standard layout
Move all content into a single LaTeX file for simplicity.
Remove all unused auxilliary files.
Rename principle file to cwp-wg-training-careers.tex.
Ignore PDF output file in git to avoid accidental checkin.

@@

Both career paths suffer in academic terms from few or poorly cited publications, few opportunities to win grants in their own right and little opportunity for impact with commerce and wider society. These three failings make an academic appointment challenging or disfavoured compared to other applicants. However, there are measures that an be taken to address all three weaknesses. 
%This makes the development of specific solutions even more difficult and cumbersome.

@@

\begin{itemize}[itemsep=0pt,parsep=0pt,topsep=5pt,partopsep=0pt]
   \item \emph{Beginners}: New collaborators with no knowledge of the tools or techniques they are expected to use. These might be undergraduates but they could also be senior researchers who have not yet or not recently engaged in software analysis. These are people in need of some kind of formal training in modern computing techniques (basically compiled and scripting languages together with operating system basics such as filesystems, version control, and command-line shells).
   \item \emph{Intermediate users}: HEP analyzers with some experience in either analysis concepts and/or tools, but  looking to supplement their experiences with more recent and modern approaches. 
   \item \emph{Advanced users}: HEP analysis experts who have mastered current technologies and implementations (at a certain point in time) and want to stay up-to-date with new advanced developments.
   \item \emph{Software developers}: HEP scientists in charge of software development in areas not limited only to analysis, such as DAQ systems, computing infrastructure, databases, pattern recongition, and complete frameworks.
\end{itemize}

@@

\begin{itemize}[itemsep=0pt,parsep=0pt,topsep=5pt,partopsep=0pt]
   \item Basic and Advanced Programming Concepts
   \begin{itemize}[itemsep=0pt,parsep=0pt,topsep=5pt,partopsep=0pt]
      \item Object oriented paradigm
      \item Compiled langauges (C++)
      \item Scripting languages (Python, Javascript,...)
   \end{itemize}
   \item Algorithmics 
   \begin{itemize}[itemsep=0pt,parsep=0pt,topsep=5pt,partopsep=0pt]
      \item Boost library
      \item STL algorithms for containers
      \item R and/or ROOT
   \end{itemize}
   \item Existing frameworks (development or application level)
   \begin{itemize}[itemsep=0pt,parsep=0pt,topsep=5pt,partopsep=0pt]
      \item Qt
      \item ROOT
      \item experiment specific framework (possibily if of potential interest outside the native experiment)
   \end{itemize}
   \item Code design (design patterns)
   \item Development tools
   \begin{itemize}[itemsep=0pt,parsep=0pt,topsep=5pt,partopsep=0pt]
      \item IDEs (Integrated Development Environment)
      \item Debuggers
      \item Profilers
   \end{itemize}
   \item Evaluation metrics
   \item “Trust” metrics such as data driven tests
   \item Specific software implementation training
   \item Good practices 
   \item Code style and clarity
   \item Scripting and data cleaning
   \item Reporting results reproducibly
   \item Writing Documentation
\end{itemize}

@@

Difficulties that have emerged in the past with respect to implementing training courses are the lack of funding and the lack of available time by experts in the field. People with enough expertise or insight in the field usually don't have time to devote to prolonged periods of student training, and, even when they can find time, the cost of setting up a training course in an effective way is often beyond what's made available by funding agencies (funds for travel, hosting, setting up a room with a computing infrastructure to allow interactive hands-on session, etc.) A possible solution is a completely different approach to training (but complementary to the already existing and successful classical efforts such as the CERN School of Computing's Bertinoro and Kit ones): instead of directly teaching to students, trainees could make use of a web-based platform to provide training materials to students. This complementary approach has several advantages over traditional ones:
\begin{itemize}[itemsep=0pt,parsep=0pt,topsep=5pt,partopsep=0pt]
   \item Tutors can add material to the web site at a very slow pace (whenever they find time to do it, one slide a week or a chapter a day)
   \item Their material, made public on a web-site, can be further expanded by collaborators (also at their own pace) or, even better and more productively, by students who decide to contribute new additions, examples, exercises etc. Such a collaborative effort allows more people to be exposed to training at any given time, creates a sense of community, and creates bridges between people in contiguous areas or research. Students can use the same platform to exchange their own examples, make suggestions and point out interesting concepts. In such a model, the possible contribution from others to the training material needs to be moderated and validated with appropriate policies.
   \item If complemented by the availability of remote virtual machines (possibly through a browser), students could have access to examples and exercises that are already embedded in their own natural environment: all the necessary tools/libraries needed to implement the exercise will be already available in the virtual machine (possibly a Docker container). With just a web browser, students could run complex examples from home, taking advantage of a remote facility that provides some storage and computing power. Important here is the concept of ``environment'': a Docker container could be set up in such a way that students will work in an exact replica of the environment they will be exposed to in their experiment (or collaboration at large). Moreover, students could be provided with Docker containers that preserves their modified environment across sessions, allowing them to develop their skills over a prolonged period of time by accessing all the files that were made persistent day by day during the training.
   \item There would no longer be a need to find the resources to host a school and pay the tutor(s) (and eventually subsidize the students to participate in training in a remote location). Students could follow the training at their own pace from wherever they happen to be. A traditional school only lasts for 5 days (10 at most) and it is difficult to cover a  subject to any significant depth in such a short time. The web approach, instead, would allow for very long and in-depth coverage of any kind of subject, and in this sense it could be a {\it complementary} approach to a traditional school. Of particular interest could be courses such as, e.g., ``Machine Learning'', ``Statistical Analyis with ROOT'', or even just ``Good practices in C++'' or ``Python Programming for scientific computing''. 
   \item Finally, this approach could allow the creation of {\it browsable} repositories of all training materials, grouped by argument, by relevance, by experiment or whatever other critieria. Students from all over the world could be exposed to a large repository of examples, exercises and in general training material from their own home.

@@

%Anecdotally, these are viewed as a great success by the CMS community in preparing new collaborators and might serve as a model for future efforts in ML training.

@@

\section*{Aknowledgements}
A.P. acknowledges support by SNF under contract 168169.

>>>>

% Copyright (C) 2018, The HSF Community White Paper authors, licence CC-BY-4.0.

% JHEP preprint template
\documentclass[12pt,a4paper]{article}
\usepackage{jheppub}

\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{xspace}

\usepackage[utf8]{inputenc}

\usepackage[style=numeric-comp,sorting=none]{biblatex}
\addbibresource{../../HSF-CWP-2017-01_roadmap/latex/cwp.bib}
\addbibresource{../../HSF-CWP-2017-01_roadmap/latex/cwp-chapters.bib}

\abstract{The rapidly increasing evolution of technology and the parallel increasing complexity of algorithmic analysis in HEP (and in science in general) require a much larger portfolio of specific skills by the developers with respect to the past. Young researchers, graduating from universities worldwide, currently do not receive adequate preparation in the very diverse fields of modern computing to respond to growing necessities posed by the most advanced experimental challenges. The vast majority of the HEP community strongly agrees on the need for an overarching umbrella of training programs to bring young researchers (but also more mature ones) up to date with new software technologies, such as, for example, concurrent programming and artificial intelligence. }

\begin{document}

\begin{tabular*}{\linewidth}{lc@{\extracolsep{\fill}}r@{\extracolsep{0pt}}}
 & & HSF-CWP-2017-02 \\
 & & June 6, 2018 \\ % use \date or hardwire e.g. December 15, 2017
 & & \\
\end{tabular*}
\vspace{2.0cm}

\title{HEP Software Foundation Community White Paper Working Group --
Training, Staffing and Careers}

%\linenumbers

\author[1]{Dario Berzano,}
\author[2]{Riccardo Maria Bianchi,}
\author[3]{Peter Elmer,}
\author[4]{Roger Jones,}
\author[5]{Michel Jouvin,}
\author[6]{Daniel S. Katz,}
\author[7]{Sudhir Malik,}
\author[8]{Dario Menasce,}
\author[6]{Mark Neubauer,}
\author[9,a]{Albert Puig}
\author[1]{and Graeme Stewart}

\affiliation[1]{CERN, Geneva, Switzerland}
\affiliation[2]{Department of Physics and Astronomy, University of Pittsburgh, Pittsburgh, PA, USA}
\affiliation[3]{Princeton University, Princeton, NJ, USA}
\affiliation[4]{Lancaster University, Lancaster, UK}
\affiliation[5]{LAL, Université Paris-Sud and CNRS/IN2P3, Orsay, France}
\affiliation[6]{University of Illinois Urbana-Champaign, Urbana, IL, USA}
\affiliation[7]{University of Puerto Rico, Mayaguez, PR, USA}
\affiliation[8]{INFN Sezione di Milano-Bicocca, Italy}
\affiliation[9]{Physik-Institut, Universität Zürich, Zürich, Switzerland}
\affiliation[a]{Supported by SNF under contract 168169.}

\maketitle

\tableofcontents
\clearpage

%-------------------------------------------------------------------------
%\newpage
%\section{Training the community  \textcolor{pink}{Will move into a NEW chapter} }
%\label{sec:training}

@@

Both career paths suffer in academic terms from few or poorly cited
publications, few opportunities to win grants in their own right and little
opportunity for impact with commerce and wider society. These three failings
make an academic appointment challenging or disfavoured compared to other
applicants. However, there are measures that an be taken to address all three
weaknesses.

% This makes the development of specific solutions even more difficult and
% cumbersome.

@@

\begin{itemize}
   \item \emph{Beginners}: New collaborators with no knowledge of the 
   tools or techniques they are expected to use. These might be undergraduates 
   but they could also be senior researchers who have not yet or not recently 
   engaged in software analysis. These are people in need of some kind of formal 
   training in modern computing techniques (basically compiled and scripting 
   languages together with operating system basics such as filesystems, version 
   control, and command-line shells).
   \item \emph{Intermediate users}: HEP analyzers with some experience in 
   either analysis concepts and/or tools, but  looking to supplement their 
   experiences with more recent and modern approaches. 
   \item \emph{Advanced users}: HEP analysis experts who have mastered current 
   technologies and implementations (at a certain point in time) and want to 
   stay up-to-date with new advanced developments.
   \item \emph{Software developers}: HEP scientists in charge of software 
   development in areas not limited only to analysis, such as DAQ systems, 
   computing infrastructure, databases, pattern recongition, and complete 
   frameworks.
\end{itemize}

@@

\begin{itemize}
   \item Basic and Advanced Programming Concepts
   \begin{itemize}
      \item Object oriented paradigm
      \item Compiled langauges (C++)
      \item Scripting languages (Python, Javascript,...)
   \end{itemize}
   \item Algorithmics 
   \begin{itemize}
      \item Boost library
      \item STL algorithms for containers
      \item R and/or ROOT
   \end{itemize}
   \item Existing frameworks (development or application level)
   \begin{itemize}
      \item Qt
      \item ROOT
      \item experiment specific framework (possibily if of potential interest outside the native experiment)
   \end{itemize}
   \item Code design (design patterns)
   \item Development tools
   \begin{itemize}
      \item IDEs (Integrated Development Environment)
      \item Debuggers
      \item Profilers
   \end{itemize}
   \item Evaluation metrics
   \item “Trust” metrics such as data driven tests
   \item Specific software implementation training
   \item Good practices 
   \item Code style and clarity
   \item Scripting and data cleaning
   \item Reporting results reproducibly
   \item Writing Documentation
\end{itemize}

@@

Difficulties that have emerged in the past with respect to implementing training
courses are the lack of funding and the lack of available time by experts in the
field. People with enough expertise or insight in the field usually don't have
time to devote to prolonged periods of student training, and, even when they can
find time, the cost of setting up a training course in an effective way is often
beyond what's made available by funding agencies (funds for travel, hosting,
setting up a room with a computing infrastructure to allow interactive hands-on
session, etc.) A possible solution is a completely different approach to
training (but complementary to the already existing and successful classical
efforts such as the CERN School of Computing's Bertinoro and Kit ones): instead
of directly teaching to students, trainees could make use of a web-based
platform to provide training materials to students. This complementary approach
has several advantages over traditional ones:
\begin{itemize}
   \item Tutors can add material to the web site at a very slow pace (whenever
   they find time to do it, one slide a week or a chapter a day) \item Their
   material, made public on a web-site, can be further expanded by collaborators
   (also at their own pace) or, even better and more productively, by students
   who decide to contribute new additions, examples, exercises etc. Such a
   collaborative effort allows more people to be exposed to training at any
   given time, creates a sense of community, and creates bridges between people
   in contiguous areas or research. Students can use the same platform to
   exchange their own examples, make suggestions and point out interesting
   concepts. In such a model, the possible contribution from others to the
   training material needs to be moderated and validated with appropriate
   policies.
   \item If complemented by the availability of remote virtual machines
   (possibly through a browser), students could have access to examples and
   exercises that are already embedded in their own natural environment: all the
   necessary tools/libraries needed to implement the exercise will be already
   available in the virtual machine (possibly a Docker container). With just a
   web browser, students could run complex examples from home, taking advantage
   of a remote facility that provides some storage and computing power.
   Important here is the concept of ``environment'': a Docker container could be
   set up in such a way that students will work in an exact replica of the
   environment they will be exposed to in their experiment (or collaboration at
   large). Moreover, students could be provided with Docker containers that
   preserves their modified environment across sessions, allowing them to
   develop their skills over a prolonged period of time by accessing all the
   files that were made persistent day by day during the training.
   \item There would no longer be a need to find the resources to host a school
   and pay the tutor(s) (and eventually subsidize the students to participate in
   training in a remote location). Students could follow the training at their
   own pace from wherever they happen to be. A traditional school only lasts for
   5 days (10 at most) and it is difficult to cover a  subject to any
   significant depth in such a short time. The web approach, instead, would
   allow for very long and in-depth coverage of any kind of subject, and in this
   sense it could be a {\it complementary} approach to a traditional school. Of
   particular interest could be courses such as, e.g., ``Machine Learning'',
   ``Statistical Analyis with ROOT'', or even just ``Good practices in C++'' or
   ``Python Programming for scientific computing''.
   \item Finally, this approach could allow the creation of {\it browsable}
   repositories of all training materials, grouped by argument, by relevance, by
   experiment or whatever other critieria. Students from all over the world
   could be exposed to a large repository of examples, exercises and in general
   training material from their own home.

@@

% Anecdotally, these are viewed as a great success by the CMS community in
% preparing new collaborators and might serve as a model for future efforts in
% ML training.

@@

\sloppy
\raggedright
\clearpage
\printbibliography[title={References},heading=bibintoc]

\printbibliography[title={references}]

\end{document}
